Presentation Title

### 2017-06-06 thesis/chapters/ch2-model.tex
Writing TX protocol

\subsection{Transaction Protocol}
The TX protocol~\footnote{It is also known as True Halves.
See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
consist of two messages, \texttt{tx\_req} and \texttt{tx\_resp}.
When a node  wishes to create a transaction,
it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
It then sends a \texttt{tx\_req} message containing the TX block to the counterparty.
Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
and sends it back to the sender in a \texttt{tx\_resp} message.
At the end of the protocol, both parties should have both versions of the TX block.

\subsection{Validation Protocol}
The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
This is the goal of the validation protocol.

It also has two message---\texttt{vd\_req} and \texttt{vd\_resp}.
When a node wishes to validate a transaction,
it sends a \texttt{vd\_req} with the sequence number $s$ to the party that created the TX block of the transaction.
Upon receiving a \text{vd\_req}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
An agreed CP block means that it has reached consensus.
A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
With the fragment, the node replies with a \texttt{vd\_resp} message.
Upon receiving a \texttt{vd\_resp}, the node checks that the fragment is created correctly.
Loosely, that means the agreed CP blocks have actually reached consensus,
the fragment has the correct hash pointers,
and TX block containing $s$ exists and is valid.

@@

The primary data structure used in our system is Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : k \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h = |B_u|$.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of TX blocks of $u$ and $C_u$ is the set of CP blocks of $u$,
then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
where $\tau$ represents the TX type and $\gamma$ represents the CP type.

@@

\subsection{Bootstrap Phase}
Recall that facilitators are computed from the consensus result,
but the consensus result is agreed by the facilitators.
Thus we have a dependency cycle.
The goal of the bootstrap phase is to give us a starting point for the cycle.

@@

\subsection{Consensus Phase}
The consensus phase begins after the new facilitators are selected.
Assuming the facilitators are of round $r$,
the goal is to agree on some $\C_r$ between the facilitators.
There are two scenarios in the consensus phase.
First is when the node is a facilitator, the other is when the node is not.
If the node is a facilitator, it must satisfy the following conditions.
\begin{enumerate}
\item \emph{Agreement}: If any correct node outputs a CP block $c$, then every
  correct node outputs $c$.
\item \emph{Total Order}: If one correct node outputs the sequence of CP blocks
  $\{c_1, c_i, \dots, c_n\}$ and another has output $\{c'_0, c'_1, \dots,
  c'_{n'} \}$, then $c_i = c'_i$ for $i \le \min(n, n')$.
\item \emph{Liveness}: All $n - t$ correct nodes terminate eventually.
\end{enumerate}

@@

\begin{algorithm}

    \caption{Function $\textsf{tx\_init}(pk_v, m)$ generates a new TX block,
    appends it to the caller $u$'s chain and sends it to the counterparty.
    It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.
    The necessary arguments are the public key of the counterparty $pk_v$ and the transaction message $m$.}
    \label{alg:tx-init}

    \begin{algorithmic}

    \State $h \gets |B_u|$
    \State $txid \in_R \{0, 1\}^{256}$ \Comment A uniformly distributed 256 bit random value.
    \State $b_{u, h} \gets \langle \textsf{H}(b_{u, h - 1}), h, txid, pk_v, m, sig_u \rangle$
    \State $B_u \gets B_u \cup \{ b_{u, h} \}$

    \State \textbf{Send} $\langle \texttt{tx\_req}, b_{u,h} \rangle$ to $v$

\end{algorithmic}
\end{algorithm}

@@

    \begin{algorithmic}
        \Function{Create}{$z$}
        \State x
        \EndFunction

    \end{algorithmic}

\end{algorithm}

>>>>

\subsection{Transaction and Validation}
The TX protocol is a simple request and response protocol.
The nodes exchange one round of messages and create new TX blocks on their respective chains.
Thus, as we mentioned before, one transaction should result in two TX blocks.

The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
Thus we need a validation protocol to counteract malicious behaviour.
When a node wish to validate one of its TX, it asks the counterparty for the \emph{fragment} of the TX.
A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX.
Upon the counterparty's response, the node checks that the CP blocks are in consensus, the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because hash collision is hard)
 to create a different chain that begins and ends with the same two CP blocks but with a different middle section.

% The TX protocol~\footnote{It is also known as True Halves.
% See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
% consist of two messages, \texttt{tx\_req} and \texttt{tx\_resp}.
% When a node  wishes to create a transaction,
% it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
% It then sends a \texttt{tx\_req} message containing the TX block to the counterparty.
% Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
% and sends it back to the sender in a \texttt{tx\_resp} message.
% At the end of the protocol, both parties should have both versions of the TX block.

% \subsection{Validation Protocol}
% The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
% This is the goal of the validation protocol.

% It also has two message---\texttt{vd\_req} and \texttt{vd\_resp}.
% When a node wishes to validate a transaction,
% it sends a \texttt{vd\_req} with the sequence number $s$ to the party that created the TX block of the transaction.
% Upon receiving a \text{vd\_req}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
% An agreed CP block means that it has reached consensus.
% A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
% With the fragment, the node replies with a \texttt{vd\_resp} message.
% Upon receiving a \texttt{vd\_resp}, the node checks that the fragment is created correctly.
% Loosely, that means the agreed CP blocks have actually reached consensus,
% the fragment has the correct hash pointers,
% and TX block containing $s$ exists and is valid.

@@

The primary data structure used in our system is Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : k \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h = |B_u|$.
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of TX blocks of $u$ and $C_u$ is the set of CP blocks of $u$,
then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
where $\tau$ represents the TX type and $\gamma$ represents the CP type.

@@

\subsection{Bootstrap Phase}
\label{sec:bootstrap}
Recall that facilitators are computed from the consensus result,
but the consensus result is agreed by the facilitators.
Thus we have a dependency cycle.
The goal of the bootstrap phase is to give us a starting point for the cycle.

@@

\subsection{Consensus Phase}
The consensus phase begins after the new facilitators are selected.
Assuming the facilitators are of round $r$,
the goal is to agree on some $\C_r$ between the facilitators.
There are two scenarios in the consensus phase.
First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
Second if the node is a facilitator, it waits for some duration $\delta$ and collect messages of type \texttt{cp\_msg}.
After $\delta$ is elapsed, it begins the atomic broadcast algorithm.
In particular, the facilitators must satisfy the following conditions.
\begin{enumerate}
\item \emph{Agreement}: If any correct node outputs a CP block $c$, then every
  correct node outputs $c$.
\item \emph{Total Order}: If one correct node outputs the sequence of CP blocks
  $\{c_1, c_i, \dots, c_n\}$ and another has output $\{c'_0, c'_1, \dots,
  c'_{n'} \}$, then $c_i = c'_i$ for $i \le \min(n, n')$.
\item \emph{Liveness}: All $n - t$ correct nodes terminate eventually.
\end{enumerate}

We remark that this procedure is the same as what is described in the bootstrap phase (\Cref{sec:bootstrap}),
but the precondition and the waiting time is different.
In particular, the bootstrap stage assumed that every node initiated and sent the \texttt{cp\_msg} simultaneously,
but here we make no such assumption. Further, the bootstrap waits for duration $\Delta$ in order to collect all the CP blocks,
here we wait for some $\delta$ which is a system variable.
We make no assumption on the value of $\delta$ in relation to $\Delta$,
thus it may be the case that not all \texttt{cp\_msg} are delivered before the the facilitators begins the atomic broadcast protocol.

At the core of the consensus phase is the atomic broadcast protocol.
While using any atomic broadcast protocol will suffice,
we use a (minimal) simplification of HoneyBadgerBFT as it is designed for blockchain systems and runs in fully asynchronous networks.
The transactions in HoneyBadgerBFT are first queued in a buffer, the main consensus algorithm starts only when the buffer reaches an optimal size.
We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
The main algorithms are kept the same, namely Asynchronous Common Subset, Reliable Broadcast and Binary Agreement.
At the end of the consensus sphase, some $\C_r$ should be agreed upon.
This bring us back to the setup phase and the cycle can be started again.

@@

The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
It is also known as True Halves TODO.
Node that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.

\begin{algorithm}
    \caption{Function $\textsf{new\_tx}(pk_v, m, txid)$ generates a new TX block and appends it to the caller $u$'s chain.
    It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.
    The necessary arguments are the public key of the counterparty $pk_v$, the transaction message $m$ and the transaction identifier $txid$.}
    \label{alg:new-tx}

    \begin{algorithmic}
    \State $h \gets |B_u|$
    \State $t_{u, h} \gets \langle \textsf{H}(b_{u, h - 1}), h, txid, pk_v, m, sig_u \rangle$
    \State $B_u \gets B_u \cup \{ t_{u, h} \}$
    \end{algorithmic}
\end{algorithm}

@@

    \begin{algorithmic}
        \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$
        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
        \State $\textsf{new\_tx}(pk_u, m, txid)$
        \State store $t_{v, j}$ as the other half of $t_{u, h}$
        \State send $\langle \texttt{tx\_reqp}, t_{u, h} \rangle$ to $v$

        \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$
        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
        \State store $t_{v, j}$ as the other half of the TX with identifier $txid$
    \end{algorithmic}
\end{algorithm}

A key feature of the TX protocol is that it is non-blocking.
TODO
